---
title: "Chapter 13 Problems"
author: "Jack Cunningham"
format: html
editor: visual
---

13.1 The Structure of the Federal Reserve System

1.1)

The United States had no central bank between 1836 and 1913 because two previous attempts had been made and both times the bank was not renewed after their charters expired. The threat of increased centralized power was the main argument against a possible central bank. Small businesses and farms disliked the central bank as it supervised local banks and limited their loans.

Congress decided to implement the Federal Reserve System in 1913 after multiple financial panics throughout the late 18th century and early 19th century. There was concern about the stability of the financial system and the National Monetary Commission was appointed by congress to research a possible solution.

1.2)

The Federal Reserve System was split into 12 districts to avoid a primary critique of previous national banks, the centralization of power. With 12 districts and member banks holding shares the idea was that each district could represent the interests of their region.

1.3)

The Board of Governors is the more important body. Since the board of governors makes up 7 of the 12 votes in the Federal Open Market Committee they have large influence over the policy decisions in their meetings. Additionally their economists present research and provide much of the data that is discussed in the FOMC meetings (the Green and Blue book).

1.4)

The Dodd-Frank act introduced new responsibilities in the Financial Stability Oversight Council, restricted the Fed's ability to make loans of last resort to specific financial institutions, class A directors (bankers) were now unable to vote for their regional bank president, more disclosures are required and the Consumer Financial Protection Bureau was established.

1.5)

a\)

The district banks are owned by the member banks of their region, in this sense the bank is private. The member banks also vote for the directors of the member banks. These district banks then make up a portion of the FOMC who vote and implement monetary policy.

b\)

The Federal Reserve System was a creation by congress and thus is suspect to congressional authority. If congress desired they could implement new rules, change existing operations or remove the Federal Reserve System altogether.

At the same time the Federal Reserve System is independent of Government in ways different than other agencies. The board of governors serve 14 year terms that have an expiration every two years to restrict the President from having power over the entire board. The Federal Reserve brings in profit which allows it to make its own budgetary decisions. The decisions and operations of the Fed are decided through the board of governors and the member bank presidents.

1.6)

a\)

Congress set up a system with inherent tension between the Federal Reserve Banks and Federal Reserve Board in an attempt to limit the centralization of power in the system. They wanted the reserve banks to have influence over policy that reflected their regions issues and beliefs.

b\)

The tension was addressed in 1935 when the Board of Governors were given 7 of 12 votes in the FOMC and the authority to set reserve requirements. In many respects the Board of Governors have the dominant power over the system. They supervise the Board Staff and have a majority of the vote in the FOMC. Particularly the chair of the Board of Governors has had a lot of freedom to pursue policy, notably Ben Bernanke in the financial crisis of 2007-2009.

1.7)

a\) A main argument for Federal Reserve Banks acting independently is their ability to react to regional problems and the reduction in centralization of the system.

An argument against the Federal Reserve Banks acting independently is that one bank could encourage growth in their region at the expense of the country's financial situation.

b\) It would not be possible for a Federal Reserve Bank to act as the New York Fed did in 1929. This is due to the fact that the board of governors and the FOMC have the authority to dictate policy not any independent district bank.

1.8)

Congress should keep the general principle of forming districts that consist of multiple industries. They should adjust naturally based on migration that has happened since 1914 (i.e. San Francisco seeing massive population growth). The boundaries would probably change due to an expansion of industry in the west. Currently the bank of San Francisco covers nearly all of the west coast, a bank of Seattle could help divide that area. In today's world the district doesn't matter as much due to the onset of digital banking. In 1914 the physical distance between banks and the regional banks mattered much more.

1.9)

```{r}
interest_paid <- 0.02
reserve_req <- .1
interest_received_fed <- 0.005
interest_earned_loan <- 0.07
```

a\)

```{r}
amount_earned <- 1000*(1-reserve_req)*interest_earned_loan + 1000*reserve_req*interest_received_fed

theor_amount_earned <- 1000*interest_earned_loan
```

The amount earned from 1,000 dollars in deposits is 63.5 dollars, the amount that would be earned without reserve requirements is 70 dollars.

b\)

In a period of high inflation rates will creep higher, that will create a greater opportunity cost for having required reserves that are only earning a minimal amount at the Fed.

1.10)

The Sunshine act is supposed to provide more transparency for when government employees meet, this helps the media properly report on what they are doing and gives the public assurance that there aren't secret backroom meetings taking place. Ben Bernanke went around those rules in 2007-2009. The question here is whether the ends justify the means. The financial crisis could have spiraled out of control but the actions of Ben Bernanke and the "four musketeers" helped calm the waters. But I still think an intentional manipulation to break the spirit of the act was not a good move. To this day there is a mistrust in the financial system and the Fed and to some extent that is due to the lack of transparency around how the Fed acted in that crisis.

1.11)

a\)

The hybrid governance model refers to the quasi government situation the reserve banks are in. They are private companies but their mandate comes from the government.

b\)

If Congress and the President were in control of the monetary policy there is good reason to believe they would act in their own interest and try to keep an economy pumping throughout their terms for political gain. This could lead to inflation that spirals out of control as no government entity wants to be the one to induce unemployment and lowered growth by raising rates.

1.12)

a\)

Class A directors are the bankers voted to serve as directors in a district bank.

b\)

The bankers would likely support keeping rates low as to reduce the banking industries opportunity cost with required reserves.

c\)

Congress established this structure to give local banks a say over policy. One of the major failings of the previous two national banks were that banks and small businesses disliked the fact they were getting their operations. This structure allowed those complaints and suggestions to be heard in a constructive manner.

13.2 How the Fed Operates

2.1)

The Fed was created by Congress, so Congress has the ability to change or even abolish the organization. Thus they are under political pressure to act in a way that does not agitate their members.

2.2)

The public interest view is that the Federal Reserve System acts in a way to benefit the general public. The goals that both groups share is price stability, high employment and economic growth.

The principal agent view is that the Federal Reserve System acts in a way to achieve as much power and independence it can and will stretch the rules to get there.

The political business cycle is a theory that the Federal Reserve will reduce rates to energize an economy before an election as to gain favor with the incumbency.
